---
title: "Arrests"
author: "*Boadiwaa*"
date: "Created on `r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(knitr)
library(dplyr)
library(ggplot2)
library(carData)
```

###  **Overview of Report**

The Arrests dataset is part of the carData package. The description from the official document accompanying the package reads: "Data on police treatment of individuals arrested in Toronto for simple possession of small quantities
of marijuana. The data are part of a larger data set featured in a series of articles in the Toronto Star
newspaper." This report seeks to derive insights according to the following Objectives:

- Black vrs White Releases
- Yearly trend of arrests
- Employment status and its relation with Release status
- Age categories for arrests
- Female vrs Male Releases
- Citizen vrs Non-citizen Releases

### Glimpse of the Source Data
```{r}
#Gives an overview of the source data for this report.
test_data <- carData::Arrests
summary(test_data)
head(test_data)
```

### Did It Matter Whether They Were Black or White?
The graph below compares the statistics on arrests and releases according to the individual's race (ie. Whether they were broadly-speaking: Black or White)

```{r, echo=FALSE, fig.cap= "Figure1: Black vrs. White Statistics" , fig.align='center'}
black_or_white <- test_data %>% group_by(released,colour) %>% summarize(Count=n())
b <- (black_or_white[4,3]/(black_or_white[4,3] + black_or_white[2,3]))*100
a <- (black_or_white[3,3]/(black_or_white[3,3] + black_or_white[1,3]))*100
percent_black_releases <- (black_or_white[3,3]/(black_or_white[3,3] + black_or_white[1,3]))*100
percent_white_releases <- (black_or_white[4,3]/(black_or_white[4,3] + black_or_white[2,3]))*100
c<- 100- percent_black_releases
d <- 100- percent_white_releases
black_or_white$percentages <- c(c,d,percent_black_releases, percent_white_releases)
ggplot(black_or_white, aes(x=released, y= percentages, fill = released)) + geom_col(colour="black") + 
  scale_fill_manual(values = c("burlywood","brown4")) + facet_wrap(~colour) + labs(title = "Black vrs White Release stats") + 
  theme(plot.title = element_text(hjust = 0.5))
```



### Yearly Trend of Arrests
```{r, echo = FALSE, fig.cap="Figure 2: Trend of Number of Arrests over the years", fig.align='center'}
arrests_trend <- test_data %>% group_by(year)%>% summarize(Count=n())
ggplot(arrests_trend, aes(x=year,y=Count)) + geom_line() +geom_point(color="brown ", size =3) + labs(title = "Yearly Trends of Arrests") +
  theme(plot.title = element_text(hjust = 0.5))
```



### Employment status and its relation with Release status
The graph below shows that, majority of those who were released were in some form of employment. Overall, those who were employed were more than those who were not employed irrespective of the outcome of the arrests.

```{r, echo = FALSE, fig.cap="Figure 3: Employment Status in relation to Arrest Outcome",fig.align='center'}
employment_factor <- test_data %>% group_by(employed,released) %>% summarize(Count=n())

ggplot(employment_factor, aes(x = employed, y=released)) + geom_point(aes(size=Count), shape=21, colour="black", fill="ivory") + 
  scale_size_area(max_size = 40, guide = "none") + geom_text(aes(label = Count), size=4)
```



### Age Categories and Number of Arrests
Majority of arrests were made in the 18 to 29 age group. Here is a visualization of the data:
```{r, echo = FALSE, fig.cap="Figure 4: Age Categories vrs Arrests", fig.align='center'}
teenagers <- test_data %>% filter(age >= 12, age < 18) %>% summarize(Count=n())
young_adults <- test_data %>% filter(age >= 18, age < 30) %>% summarize(Count=n())
prime_time <- test_data %>% filter(age >= 30, age < 46) %>% summarize(Count=n())
mid_age_to_elderly <- test_data %>% filter(age >= 46) %>% summarize(Count=n())

Number_of_Arrests <- c(teenagers[1,1],young_adults[1,1],prime_time[1,1],mid_age_to_elderly[1,1])
Age_group <- c("12 to 17", "18 to 29", "30 to 45", "46 and above")
Age_stats <- data.frame(Age_group, Number_of_Arrests)

ggplot(Age_stats, aes(x=Age_group, y= Number_of_Arrests, fill=Age_group)) + geom_col(colour="black") + 
  scale_fill_manual(values = c("coral4","burlywood","deeppink4","brown4")) + labs(title = "Age categories and Number of Arrests") + 
  theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "none")
```



### Stats of the Sexes
The table below shows the breakdown according to gender.
```{r, echo=FALSE}
male_vrs_female <-test_data %>% group_by(sex) %>% summarize(Count=n())
kable(male_vrs_female, col.names = c("Gender", "Number of Arrests"), caption = "Table 1: Male Vrs. Female Arrests")
```



### Citizen Vrs Non-Citizen

The graph below compares the number of Citizens who were arrested and released, with the corresponding statistics in non-citizens:
```{r, echo=FALSE, fig.cap="Figure 5: Citizens Vrs. Non-Citizens Release Stats", fig.align='center'}
citizen_vrs_non_citizen <- test_data %>% group_by(citizen,released) %>% summarize(Count=n())
ggplot(citizen_vrs_non_citizen, aes(x=released, y= Count, fill = citizen)) + geom_col(colour="black") + 
  scale_fill_manual(values = c("burlywood","brown4")) + facet_wrap(~citizen) + labs(title = "Citizen vrs Non-citizen Release stats") + 
  theme(plot.title = element_text(hjust = 0.5)) + geom_text(aes(label= Count), vjust=-0.2) + theme(strip.text = element_blank())

```

This brings us to the end of the brief report.

**Thank you!**
